### 2019-07-01 response.txt
Add response to reviews

Thank you for the very helpful comments!

The main feeling among the reviewers and AE was that the article did not go far enough -- that its proposals were too timid and incremental.  We took this response to heart and essentially rewrote the whole thing from top to bottom, removing most discussion of short-term mitigations and replacing it with two concrete and specific proposals for actions that ACM can and should take on a large scale -- one reasonably easy, one much more ambitious.  We have also removed the uninformative figures and completely revamped the remaining figure to make it (we hope) much more interesting.

The new version does still diverge from the recommendations of the reviewers in two significant respects:

1) Reviewer 2 felt that the main focus of the article should be on the need to move past conferences altogether.  While we agree that this is where things may need to move in the medium to long term, it is very far from where the community is now, and we feel this is for good reason: CS conferences offer a great deal of value in addition to the environmental harm that their emissions create, and it is far from clear at this point that simply deleting conferences would be a net benefit to society.  The discussion about balancing benefits and harms is going to be a long and painful one, and we do not feel ready to prejudge the outcome at this point.  What is needed is to begin the discussion, and we hope that the new version of the article helps do this.

We did add a paragraph at the end acknowledging reviewer 2's observation that ACM's business model relies heavily on conference revenue and therefore needs rethinking.

2) Reviewer 2 also strongly objected to our even mentioning carbon offsets as a potential mitigation strategy.  After much discussion among ourselves, we decided not to do this: The reviewer's position is reasonable (and held by lots of reasonable and well-informed people), but not universal, and we respectfully continue to believe (and assert in the article) that carefully chosen offsets offer a useful, short-term way to mitigate the harm we are doing right now, buying the world (and ACM) a little more time to make the deeper adjustments that we 100% agree are necessary.  However, we have greatly shrunk the discussion of offsets, and we've included a new paragraph explicitly mentioning the dangers of bogus offsets and the greater danger of encouraging complacency.  We hope this will be viewed as a reasonable compromise.